---

Author: Neal Barton
Version: 1
Date: 14/04/2020
Output: word_document
Title: Generalised Additive Model Script
output:
  word_document: default

---

## Install relevant packages

Load relevant packages from the R library. In this study we used mgcv for the gam models and mgcViz for the visualisations. extrafont, ggthemes, ggplot were used for the aesthetics of the plots, whilst gridExtra was used to set the grid matrix for the final layout.  

```{r install relevant packages, eval=FALSE, include=TRUE}

# For Generalised Additive Model
library(mgcv)

# For the plots
library(mgcViz)
library(extrafont)
library(extrafontdb)
library(ggthemes)
library(ggplot2)
library(gridExtra)

# Extend memory due to large data sets 
memory.limit(60000000)

```

## Generalized Additive Model 

In the ‘mgcv’ package we used ‘bam’ for large data sets, penalized cubic regression splines (bs = ‘cs’) and use select = TRUE which removes variables with coefficients at 0. The knots (‘k =’) is left as default, family used ‘poisson()’, the method is restricted maximum likelihood ‘REML’ and we use an offset of length to account for different pipe lengths.

```{r generalized additive model, eval=FALSE, include=TRUE}

model <- bam(nBurstsReactive ~ offset(log(length)) + Diameter_Band + Age_Band + s(Temp, bs = "cs") + s(TchgW1, bs = "cs") + s(TchgW2, bs = "cs") + 
             s(DaysAirFrost, bs = "cs") + s(SMD, bs = "cs") + s(SMDchgW1, bs = "cs") + s(SMDchgW2, bs = "cs") + s(pH, bs = "cs") +
             Shrink_Swell + Subsoil + HOST, select = TRUE, family = poisson(), data = df, method = "REML")

# Check the model summary
summary(model) 

# Check model diagnostics - to see if the residuals are significant or not
gam.check(model)

# Check for pairwise concurvity
concurvity(model, full = FALSE)

```

## Plot Visualisaiton

We use mgcViz to extract the partial plots and then extract each covariate individually to create the desired matrix of plots. Adapted here from:

Fasiolo, M., Nedellec, R., Goude, Y., Wood, S.N., 2019. Scalable Visualization Methods for Modern Generalized Additive Models. J. Comput. Graph. Stat. 1–9. https://doi.org/10.1080/10618600.2019.1629942 

```{r plot visualisaiton, eval=FALSE, include=TRUE}

# Set themes for each plot
themeSize <- theme(text = element_text(family = "serif"), 
                   axis.text = element_text(size = 11),
                   axis.title = element_text(size = 11), 
                   legend.title = element_text(size = 11), 
                   legend.text = element_text(size = 11))

# Create a figure list with all plots 
figure_1 <- list()

#  use c - for categorical terms and s for smoothed terms
figure_1[[1]] <- plot(pterm(gam_plot, 1), n = 200) + 
  l_fitPoints(size = 1, col = "red") +
  l_ciBar(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  labs(x = expression("Diameter Band (mm)"), y = expression(c * "(" * "<df>" * ")")) + 
  scale_x_discrete(labels = c("1" = "<166", "2" = "166-<321", "3" = "321-<625", "4" = ">625")) +
  theme(axis.text.x = element_text(angle = 15, hjust = 1)) +
  ggtitle(~italic("<p-value>")) +
  theme_classic() +
  themeSize

figure_1[[2]] <- plot(pterm(gam_plot, 2), n = 200) + 
  l_fitPoints(size = 1, col = "red") +
  l_ciBar(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  theme_classic() +
  labs(x = expression("Age Band"), y = expression(c * "(" * "<df>" * ")")) + 
  scale_x_discrete(labels = c("1" = "<1881",
                              "2" = "1881-<1900",
                              "3" = "1900-<1920", 
                              "4" = "1920-<1940", 
                              "5" = "1940-<1960", 
                              "6" = "1960-<1980",
                              "7" = "1980-<2000",
                              "8" = "2000-<2021")) + 
  theme(axis.text.x = element_text(angle = 15, hjust = 1)) +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[3]] <- plot(sm(gam_plot, 1), n = 200) + 
  l_fitLine(linetype = 1, colour = "red") +
  l_ciLine(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  labs(x = expression("Temperature ("*~degree*C*")"), y = expression(s * "(" * "<df>" * ")")) +
  theme_classic() +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[4]] <- plot(sm(gam_plot, 2), n = 200) + 
  l_fitLine(linetype = 1, colour = "red") +
  l_ciLine(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  labs(x = expression("Temp Change Over 1 Week ("*~degree*C*")"), y = expression(s * "(" * "<edf>" * ")")) +
  theme_classic() +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[5]] <- plot(sm(gam_plot, 3), n = 200) + 
  l_fitLine(linetype = 1, colour = "red") +
  l_ciLine(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  labs(x = expression("Temp Change Over 2 Weeks ("*~degree*C*")"), y = expression(s * "(" * "<edf>" * ")")) +
  theme_classic() +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[6]] <- plot(sm(gam_plot, 4), n = 200) +
  l_fitLine(linetype = 1, colour = "red") +
  l_ciLine(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  labs(x = expression("Days Air Frost in a Month"), y = expression(s * "(" * "<edf>" * ")")) +
  theme_classic() +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[7]] <- plot(sm(gam_plot, 5), n = 200) + 
  l_fitLine(linetype = 1, colour = "red") +
  l_ciLine(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  labs(x = expression("SMD (mm)"), y = expression(s * "(" * "<edf>" * ")")) +
  theme_classic() +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[8]] <- plot(sm(gam_plot, 6), n = 200) + 
  l_fitLine(linetype = 1, colour = "red") +
  l_ciLine(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  labs(x = expression("SMD Change Over 1 Week (mm)"), y = expression(s * "(" * "<edf>" * ")")) +
  theme_classic() +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[9]] <- plot(sm(gam_plot, 7), n = 200) + 
  l_fitLine(linetype = 1, colour = "red") +
  l_ciLine(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  labs(x = expression("SMD Change Over 2 Weeks (mm)"), y = expression(s * "(" * "<edf>" * ")")) +
  ggtitle(~italic("<p-value>")) +
  theme_classic() +
  themeSize

figure_1[[10]] <- plot(pterm(gam_plot, 4), n = 200) + 
  l_fitPoints(size = 1, col = "red") +
  l_ciBar(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  theme_classic() +
  labs(x = expression("Subsoil Texture (70 cm Depth)"), y = expression(c * "(" * "<df>" * ")")) +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[11]] <- plot(pterm(gam_plot, 3), n = 200) + 
  l_fitPoints(size = 1, col = "red") +
  l_ciBar(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  theme_classic() +
  labs(x = expression("Shrink Swell (% Vol)"), y = expression(c * "(" * "<df>" * ")")) + 
  scale_x_discrete(labels = c("1" = "<3",
                              "2" = "3-5",
                              "3" = "5-12",
                              "4" = "12-15",
                              "6" = "12-15*")) +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[12]] <- plot(sm(gam_plot, 8), n = 200) + 
  l_fitLine(linetype = 1, colour = "red") +
  l_ciLine(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  labs(x = expression("pH"), y = expression(s * "(" * "<edf>" * ")")) +
  theme_classic() +
  ggtitle(~italic("<p-value>")) +
  themeSize

figure_1[[13]] <- plot(pterm(gam_plot, 5), n = 200) + 
  l_fitPoints(size = 1, col = "red") +
  l_ciBar(linetype = 3, colour = "blue", level = 0.99) +
  l_rug() +
  theme_classic() +
  labs(x = expression("HOST"), y = expression(c * "(" * "<df>" * ")")) + 
  ggtitle(~italic("<p-value>")) +
  themeSize


# To extract the smoothed variable - Temp - Works fine
fig_layout <- matrix(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15), 
                     nrow = 5, 
                     ncol = 3, 
                     byrow = T)


png("<data_location>", units = "mm", res = 1000, height = 280, width = 210)

grid.arrange(grobs = lapply(figure_1, "[[", "ggObj"), layout_matrix = fig_layout)

dev.off()

```


