---

Author: Neal Barton
Version: 1
Date: 24/06/2021
Output: word_document
Title: Gradient Boosting Decision Tree Model
output:
  word_document: default

---

## Load packages

The gradient boosing model packages. 

```{r load packages, eval=FALSE, include=TRUE}

# Load packages

library(caret) # for developing the model
library(pROC) # for AUC calculations
library(dplyr) # model processing
library(purrr) # for functional programming (map)
library(ModelMetrics) # For looking at model metrics
library(ggplot2) # for the plots
library(ggthemes) # to improve the plots

# Increase memory

memory.limit(30000000)

```

## Run Model

This is the final model based on the tuned hyperparameters. 

```{r run model, eval=FALSE, include=TRUE}

# Load dataset
df <- readRDS(("DATA LOCATION"))

# Stratify the sampling based on a split of the class and different representative materials 
set.seed(5627)  # set the random seed for reproducibility
strat <- stratified(df, c('mat', 'class'), 0.7, bothSets = T)

# Collect training and testing data sets from the stratified sampling function
train <- as.data.frame(strat$SAMP1)
test <- as.data.frame(strat$SAMP2)

# mutate all characters into factors for model to work
train <- train %>% mutate_if(is.character, as.factor)
test <- test %>% mutate_if(is.character, as.factor)


# Set up control function for training, use 5 fold cross validation
ctrl <- trainControl(method = "cv",
                     number = 5, 
                     returnResamp = 'none',
                     summaryFunction = twoClassSummary,
                     classProbs = T,
                     savePredictions = T,
                     verboseIter = F)


# Set up trainng grid - add parameters of optimal trained model
gbmGrid <-  expand.grid(interaction.depth = 15,             
                        n.trees = 1649,                                   
                        shrinkage = 0.01,                                     
                        n.minobsinnode = 2)                  

# set the random seed for reproducibility and run model
set.seed(5627)  
model <- train(class ~ .,
             data = train,
             method = "gbm",
             metric = "ROC",
             tuneGrid = gbmGrid,
             verbose = FALSE,
             trControl = ctrl)


# Run model on test data
test$gbm <- predict(model, newdata = test, type = "prob") [, "yes"]

# Check variable importance
summary(model)

# print model summary
print(model)

```

## ROC/AUC 

Run the models ROC/AUC metrics, code adapted from: 

Martin, D.P. 2016. Handling Class Imbalance with R and Caret – An Introduction. http://dpmartin42.github.io/posts/r/imbalanced-classes-part-1

```{r ROC/AUC, eval=FALSE, include=TRUE}

# Create a model list
model_list <- list(gbm = model)

# Build functon to create predictions
test_roc <- function(model, data) {
  
  roc(data$class,
      predict(model, newdata = data, type = "prob")[, "yes"])

}
  
# Examine results for test set
model_list_roc <- model_list %>%
  map(test_roc, data = test)

model_list_roc %>%
  map(auc)

results_list_roc <- list(NA)
num_mod <- 1

for(the_roc in model_list_roc){
  
  results_list_roc[[num_mod]] <- 
    data_frame(tpr = the_roc$sensitivities,
               fpr = 1 - the_roc$specificities,
               model = names(model_list)[num_mod])
  
  num_mod <- num_mod + 1
  
}

results_df_roc <- bind_rows(results_list_roc)

# Plot ROC curve for all 5 models

ggplot(aes(x = fpr,  y = tpr, group = model), data = results_df_roc) +
  geom_line(aes(color = model), size = 0.5) +
  theme_hc() +
  scale_colour_manual(values = c("#4E78A7")) +
  theme(text = element_text("sans", size = 11),
        axis.title.x = element_text(size = 11),
        axis.text.x = element_text(size = 11),
        axis.title.y = element_text(size = 11),
        axis.text.y = element_text(size = 11),
        axis.line = element_line(size = 0.5, colour = "black"),
        legend.position = "none",
        legend.title = element_blank()) +
  ylab("True Positive Rate") +
  xlab("False Positive Rate") +
  geom_abline(intercept = 0, slope = 1, color = "grey", size = 1)

```

## Explore Thresholds

Use thresholder to explore the optimal threshold.

```{r explore thresholds, eval=FALSE, include=TRUE}

# Calcuate best thresholds
thresholder(MODEL, threshold = seq(.01, 1, by = 0.01), final = TRUE) # Set search granularity

thres <- as.data.frame(thres)

```


## Confusion matrix and MCC accuracy

calcualte confusion matrix for different thresholds and calcualte the MCC measure of accuracy for each scenario. 

```{r confusion matrix and MCC accuracy, eval=FALSE, include=TRUE}

test$opt_thres <- ifelse(test$gbm < <value>, 0, 1) # add relevant threshold value

# confusion matrix for testing accuracy of binomial model
con_mat <- ftable(test$opt_thres, test$class)

# =============== stats ============= #

TN <- as.numeric(con_mat[1])
FN <- as.numeric(con_mat[2])
FP <- as.numeric(con_mat[3])
TP <- as.numeric(con_mat[4])

# Print the values from the confusion matrix
print("TP")
TP/(TP + FP) * 100
print("FP")
FP/(FP + TP) * 100
print("FN")
FN/(TN + FN) * 100
print("TN")
TN/(TN + FN) * 100

# print the MCC
((TP * TN) - (FP * FN)) / sqrt((TP + FP) * (TP + FN) * (TN + FP) * (TN + FN))


```

