---

Author: Neal Barton
version: 1
Date: 14/04/2020
Output: word_document
Title: An Evaluation of Pipe Failure in Drinking Water Networks
output:
  word_document: default

---

## Install relevant packages

```{r Import Packages, eval=FALSE, include=TRUE}

# Install relevant packages

library(extrafont)
library(dplyr)
library(data.table)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(stringr)

```

## Poisson mean function with 95% confidence interval 

Poisson mean distribution with 95 % confidence interval: Written by Adam Howes 2017

````{r Poisson mean, eval=FALSE, include=TRUE}

knitr::opts_chunk$set(eval = FALSE, message = FALSE, warning = FALSE)

# 95% Confidence Interval Function 
  poisson_mean <- function(num, length) {
    
    log_estimate <- rep(0, length(num))
    log_low_limit <- log_estimate
    log_high_limit <- log_estimate
    for (i in 1:length(num)) {
      glm1 <- glm(num[i] ~ 1 + offset(log(length[i])), family = poisson)
      log_estimate[i] <- summary(glm1)$coefficients[,1]
      log_se <- summary(glm1)$coefficients[,2]
      log_low_limit[i] <- log_estimate[i] - (1.96*log_se)
      log_high_limit[i] <- log_estimate[i] + (1.96*log_se)
    }
    estimate <- exp(log_estimate)
    low_limit <- exp(log_low_limit)
    high_limit <- exp(log_high_limit)
    send_back <- array(c(estimate, low_limit, high_limit), dim = c(length(num), 3))
    colnames(send_back) <- c("estimate", "low_limit", "high_limit")
    return(as.data.frame(send_back))
  }


```

## Calculate seasonal failure rates for each material

Calculate seasonal failure rates for each material in the Standard Ofwat format of bursts/1000km/year.

```{r seasonal failure rate, eval=FALSE, include=TRUE}

knitr::opts_chunk$set(eval = FALSE, message = FALSE, warning = FALSE)

  # get unique pipe length (because pipe length is repeated by year which distorts final results)
  df_unique_no <- (length(unique(mat$Week_end)))
  df_unique_no <- as.numeric(df_unique_no)
  
  # group data by month
  df <- df[ , list(length = sum(length), bursts = sum(nBurstsReactive)), by = c(var)]
  
  # caculate length of pipe
  df$length <- (((df$length / 1000000) / df_unique_no))
  
  # Use the poisson mean function to caculate EV and 95% CI for Bursts per week per 1000km / years
  df <- bind_cols(df, poisson_mean(df$bursts, df$length)/year) # change year as appropriate
  
  # Reorder factor levels for ggplot
  df$Month <- factor(df$var, levels = df$var[order = c(12,1,2,3,4,5,6,7,8,9,10,11)])
  
  saveRDS(df, "FILE_LOCATION, ",(mat),"_",(var),".rds")

```

## Create plots using ggplot function

Example plot below for seasonal variation plot. All plots were tailored to each variable: season, pipe age and diameter. 

```{r plots, eval=FALSE, include=TRUE}

knitr::opts_chunk$set(eval = FALSE, message = FALSE, warning = FALSE)

ggplot(df, aes(x = Month, y = estimate, group = Material)) +
  geom_line(aes(color = Material, linetype = Material), size = 1) +
  geom_point(aes(shape = Material, color = Material), size = 2) +
  theme_classic() +
  theme(legend.text = element_text(family = "Times New Roman", size = 12),
        axis.title = element_text(family = "Times New Roman", size = 12),
        axis.text = element_text(family = "Times New Roman", size = 12),
        legend.title = element_text(family = "Times New Roman", size = 12)) +
  guides(fill = guide_legend(title = "Material")) +
  scale_x_discrete(name = "Month", labels = c("1" = "Jan", "2" = "Feb", "3" = "Mar", "4" = "Apr", "5" = "May", 
                                              "6" = "Jun", "7" = "Jul", "8" = "Aug", "9" = "Sep", 
                                              "10" = "Oct", "11" = "Nov", "12" = "Dec")) +
  labs(y = "Bursts/1000 km/Year")
  # geom_smooth((aes(group = 1)), se = FALSE, linetype = "dashed", color = "red", size = 0.5)
  
  ggsave("Output/Seasonal_Failure_Rate/All.png", dpi = 1000, width = 210, height = 148, units = 'mm')

```

